---
title: "Final project"
author: "Dariel Nunez, Juan Liberos, Patrick Isky, Dariel Jimenez"
format: html
editor: visual
---

# Data

[Data](https://catalog.data.gov/dataset/national-obesity-by-state-d765a/resource/5fd75a16-98e4-467e-ab49-156fe25de504)(Download the xlsx version)

```{r}
library(tidyverse)
library(effects)
library(readxl)

state <- read_excel("obesity_data/LakeCounty_Health_-1642538411848975661.xlsx")

```

#### Variables of the data state:

NAME(the states) = categorical

Obesity = numeric continuous

Shape_area = numeric continuous

Shape_length = numeric continuous

# Question 1(categorical vs numeric continuous)

Which state has the highest percent of obesity? Which has the lowest?

```{r}
state|>
  summarize(Highest= max(Obesity),
            Lowest = min(Obesity))
```

### Plot

```{r}
state|>
   ggplot(aes(y = reorder(NAME, Obesity), x = Obesity))+
  geom_point()
```

### Conclusion

The state with the highest obesity rate is Lousiana with a 36.2% of obesity. The state with the lowest obesity rate is colorado with a 20.2%.

# Question 2 (categorical vs numeric continuous)

Which region has on average the highest obesity percentage?

### Data

```{r}
northeast <- c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "New Jersey", "New York", "Pennsylvania", "Rhode Island", "Vermont")

midwest <- c("Illinois", "Indiana", "Iowa", "Kansas", "Michigan", "Minnesota", "Missouri", "Nebraska", "North Dakota", "Ohio", "South Dakota", "Wisconsin")

south <- c("Alabama", "Arkansas", "Delaware", "Florida", "Georgia", "Kentucky", "Louisiana", "Maryland", "Mississippi", "North Carolina", "Oklahoma", "South Carolina", "Tennessee", "Texas", "Virginia", "West Virginia", "Puerto Rico")

west <- c("Alaska", "Arizona", "California", "Colorado", "Hawaii", "Idaho", "Montana", "Nevada", "New Mexico", "Oregon", "Utah", "Washington", "Wyoming")



by_region<- state|>
  mutate (region = ifelse(NAME %in% northeast, "northeast", 
                          ifelse(NAME %in% midwest, "midwest",
                                 ifelse(NAME %in% south, "south",
                                        ifelse(NAME %in% west, "west", "District of Columbia")))))
```

#### Variables of the data by_region:

NAME(the states) = categorical

Obesity = numeric continuous

region = categorical

Shape_area = numeric continuous

Shape_length = numeric continuous

```{r}
by_region|>
  group_by(region)|>
  summarize(mean = mean(Obesity), median = median(Obesity))
```

### Linear Regression

```{r}
model <- lm(Obesity~region, data = by_region)

summary(model)
```

The P value is less than 0.05 so we can reject the null hypothesis. The variance is 0.5259, meaning that the model explains 52.59% of the variation.

### Plot

```{r}
by_region|>
  group_by(region)|>
  summarize(mean = mean(Obesity))|>
  ggplot(aes(x = mean, y = reorder(region, mean)))+
  geom_col()
```

### Conclusion

The region with the highest obesity rate is the South with 32.28%. The region with the lowest rate of obesity is the West with 26.38%. The District of Columbia has a obesity rate of 22.10%.
